D3064R0: OOTA Execution is Provably Vacuous

### 2023-12-11 ootavacuous.tex
Peter Sewell December 10 feedback
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

@@

@@

\begin{table}
\centering
\begin{tabular}{c|c|l}
Actual Code	& Implementation	& Result \\
\hline
sdep		& sdep			& Temporal \\
\cline{2-3}
		& $\neg$sdep		& BUG!!! \\
\hline
$\neg$sdep	& sdep			& Temporal \\
\cline{2-3}
		& $\neg$sdep		& Atemporal \\
\end{tabular}
\caption{Semantic-Dependency Classification}
\label{tab:Semantic-Dependency Classification}
\end{table}

@@

>>>>

\subsubsection{Semantic Dependencies and I/O}
\label{sec:Semantic Dependencies and I/O}

A computation might include I/O:

\begin{quote}
	\co{x = y * input_int();}
\end{quote}

Then there is a semantic dependency from \co{y} to \co{x} only if
the number input is non-zero.
Similar issues arise when reading timestamps.

In addition, I/O operations are observed behavior, which greatly restricts
optimizations involving them.

@@

A special case is as follows:

\begin{quote}
	\co{x = y & 0x1 ? do_odd(y) : do_even(y);}
\end{quote}

The key point here is that diferent values might cause almost completely
disjoint code to be executed.

@@

\begin{table}
\centering
\begin{tabular}{c|c|l}
Actual Code	& Implementation	& Result \\
\hline
sdep		& sdep			& Temporal \\
\cline{2-3}
		& $\neg$sdep		& Fix C++ bug! \\
\hline
$\neg$sdep	& sdep			& Temporal \\
\cline{2-3}
		& $\neg$sdep		& Atemporal \\
\end{tabular}
\caption{Semantic-Dependency Classification}
\label{tab:Semantic-Dependency Classification}
\end{table}

@@

% Objections:
% "Atomic operations are not observable behavior."
%	They are for volatile atomic objects.  Plus an implementation
%	whose analysis is confined to a thread would need to assume
%	that any atomic store might lead to observable behavior in some
%	other thread.  An implementation with a global view must preserve
%	observable behavior regardless of OOTA cycles.	Plus if an OOTA
%	cycle does not affect observable behavior, who cares?
% "The C++ abstract machine is independent of the laws of physics."
%	That might be, but any real implementation of that abstract
%	machine will be governed by those laws.
% "Real hardware can exhibit OOTA, as exemplified by audio feedback."
%	Such hardware does not run C++.  Also, audio feedback is more
%	closely approximated by a loop calculating sound intensity than
%	by OOTA, especially given that you must include the audio
%	properties of the enclosing space in analyzing the feedback loop.
